### 2018-05-02 doc/paper.tex
Adding two lemmas from JLCS10 about the error bound.
Also

- Making numbering of most things (everything except footnotes)
  be section specific
- Making vdots be horizontally aligned with equals sign
- Adding a footnote about the change of notation for the terms
  generated by the de Casteljau algorithm
- Removing a theorem in the "review" (Section 3) because it is
  now in subsection 4.2 (where I am attempting to prove / compute
  an a priori error bound)

\usepackage{amsthm,amssymb,amsmath}
%% H/T: https://tex.stackexchange.com/a/202168/32270
\usepackage{textcomp}
%% H/T: https://tex.stackexchange.com/a/56877/32270
\usepackage{algorithm}
\usepackage{algpseudocode}
%% H/T: https://tex.stackexchange.com/a/244805
\usepackage{booktabs,siunitx}
%% H/T: https://tex.stackexchange.com/a/4217/32270
\usepackage{mathtools}
%% H/T: https://tex.stackexchange.com/a/33869/32270
\makeatletter
\newenvironment{breakablealgorithm}
  {% \begin{breakablealgorithm}
   \begin{center}
     \refstepcounter{algorithm}% New algorithm
     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
     \renewcommand{\caption}[2][\relax]{% Make a new \caption
       {\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}%
       \ifx\relax##1\relax % #1 is \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
       \else % #1 is not \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
       \fi
       \kern2pt\hrule\kern2pt
     }
  }{% \end{breakablealgorithm}
     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
   \end{center}
  }
\makeatother
%% H/T: https://tex.stackexchange.com/a/202047/32270
%%      https://tex.stackexchange.com/a/32463/32270
\usepackage[labelfont=bf]{caption}
%% H/T: https://tex.stackexchange.com/a/71153/32270
\usepackage[nottoc,notlot,notlof]{tocbibind}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}

@@

Next, we recall the de Casteljau algorithm:

@@

The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
\cite{Mainar1999}, \cite{Farouki1987}):
\begin{equation}
\cond{p, s} = \frac{\widetilde{p}\left(s\right)}{
  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
  \left|p(s)\right|},
\end{equation}
where \(B_{j, n} \geq 0\).

@@

\noindent  When comparing this computed error to the exact error, the
difference depends only on \(s\) and the Bernstein
coefficients of \(p\):

\begin{theorem}[\cite{Jiang2010}]
  If no underflow occurs and \(n \geq 2\), then
  \begin{equation}
    \left|\db{1}_0^{(0)} - \cdb{1}_0^{(0)}\right| \leq
    2 \gamma_{3n + 1} \gamma_{3(n - 1)}
    \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s) =
    2 \gamma_{3n + 1} \gamma_{3(n - 1)} \widetilde{p}(s).
  \end{equation}
\end{theorem}

\noindent Using the bound on the error in \(\db{1}\), the algorithm can
be shown to be as accurate as if the computations were done in twice
the working precision:

@@

In order to raise from twice the working precision to \(K\) times the
working precision, we continue using EFTs when computing
\(\cdb{1}^{(k)}\). By tracking the round-off from each
floating point evaluation via an EFT, we can form a cascade of global errors:
\begin{align}
  b_j^{(k)} &= \widehat{b}_j^{(k)} + \db{1}_j^{(k)} \\
  \db{1}_j^{(k)} &= \cdb{1}_j^{(k)} + \db{2}_j^{(k)} \\
  \db{2}_j^{(k)} &= \cdb{2}_j^{(k)} +
  \db{3}_j^{(k)} \\
  & \vdots \nonumber
\end{align}
In the same way local error can be tracked when updating
\(\widehat{b}_j^{(k)}\), it can be tracked for updates that happen down
the cascade:
\begin{alignat}{4}
  (1 - s) \cdot \widehat{b}_j^{(k + 1)} &+
  s \cdot \widehat{b}_{j + 1}^{(k + 1)} &&  &&=
  \widehat{b}_j^{(k)} &&+ \ell_{1, j}^{(k)} \\
  (1 - s) \cdot \cdb{1}_j^{(k + 1)} &+
  s \cdot \cdb{1}_{j + 1}^{(k + 1)} &&+ \ell_{1, j}^{(k)} &&=
  \cdb{1}_j^{(k)} &&+ \ell_{2, j}^{(k)} \\
  (1 - s) \cdot \cdb{2}_j^{(k + 1)} &+
  s \cdot \cdb{2}_{j + 1}^{(k + 1)} &&+ \ell_{2, j}^{(k)} &&=
  \cdb{2}_j^{(k)} &&+ \ell_{3, j}^{(k)} \\
  &  &&  &&\vdots && \nonumber
\end{alignat}

@@

Noting that \(\ell_{F, j}\) contains \(5F - 2\) terms, one can
show that Algorithm~\ref{alg:k-comp-de-casteljau} requires
\begin{equation}
(15K^2 + 11K - 34)T_n + K + 5 =
\mathcal{O}\left(n^2 K^2\right)
\end{equation}
flops to evaluate a degree \(n\) polynomial, where \(T_n\) is the
\(n\)th triangular number. As a comparison, the non-compensated form of
de Casteljau requires \(3 T_n + 1\) flops. In total this will require
\((3K - 4)T_n\) uses of \texttt{TwoProd}. On hardware that supports
FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
instead, lowering the flop count by \(15(3K - 4)T_n\).

@@

@@

Figure~\ref{fig:compensated-k} shows the relative forward errors compared
against the condition number. To compute relative errors, each input and
coefficient is converted to a fraction (i.e. infinite precision) and then
compared to the corresponding computed values. Similar tools are used to
\textbf{exactly} compute the condition number, though here we can rely
on the fact that \(\widetilde{p}\left(s\right) = (s - 1)
\left(\frac{s}{2} - \frac{3}{4}\right)^7\). Once the relative errors and
condition numbers are computed as fractions, they are rounded to the
nearest IEEE-754 double precision value. As in \cite{Jiang2010}, we use
values \(\left\{\frac{3}{4} - (1.3)^j\right\}_{j=-5}^{-90}\)\footnote{As with
\(0.1\), it's worth noting that \((1.3)^j\) can't be represented exactly in
IEEE-754 double precision. However, this geometric series still serves a
useful purpose since it continues to raise \(\cond{p, s}\) as \(j\) decreases
away from \(0\) and because it results in ``random'' changes in the bits of
\(0.75\) that are impacted by subtracting \((1.3)^j\).}. The curves for
\texttt{DeCasteljau} and \texttt{CompDeCasteljau} trace the same paths seen
in \cite{Jiang2010}. In particular, \texttt{CompDeCasteljau} has a relative
error that is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
\(1 / \mach\), at which point the relative error increases linearly with
the condition number until it becomes \(\mathcal{O}(1)\) when
\(\cond{p, s}\) reaches \(1 / \mach^2\).
Similarly, the relative error in \texttt{CompDeCasteljau3}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 3\))
is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
\(1 / \mach^2\) at which point the relative error increases linearly
to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^3\)
and the relative error in \texttt{CompDeCasteljau4}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 4\))
is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
\(1 / \mach^3\) at which point the relative error increases linearly
to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^4\).

>>>>

\usepackage{amsthm,amssymb,amsmath}
%% H/T: https://tex.stackexchange.com/a/202168/32270
\usepackage{textcomp}
%% H/T: https://tex.stackexchange.com/a/56877/32270
\usepackage{algorithm}
\usepackage{algpseudocode}
%% H/T: https://tex.stackexchange.com/a/244805
\usepackage{booktabs,siunitx}
%% H/T: https://tex.stackexchange.com/a/4217/32270
\usepackage{mathtools}
%% H/T: https://tex.stackexchange.com/a/33869/32270
\makeatletter
\newenvironment{breakablealgorithm}
  {% \begin{breakablealgorithm}
   \begin{center}
     \refstepcounter{algorithm}% New algorithm
     \hrule height.8pt depth0pt \kern2pt% \@fs@pre for \@fs@ruled
     \renewcommand{\caption}[2][\relax]{% Make a new \caption
       {\raggedright\textbf{\ALG@name~\thealgorithm} ##2\par}%
       \ifx\relax##1\relax % #1 is \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##2}%
       \else % #1 is not \relax
         \addcontentsline{loa}{algorithm}{\protect\numberline{\thealgorithm}##1}%
       \fi
       \kern2pt\hrule\kern2pt
     }
  }{% \end{breakablealgorithm}
     \kern2pt\hrule\relax% \@fs@post for \@fs@ruled
   \end{center}
  }
\makeatother
%% H/T: https://tex.stackexchange.com/a/28334/32270
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\counterwithin{equation}{section}
\counterwithin{algorithm}{section}
%% H/T: https://tex.stackexchange.com/a/202047/32270
%%      https://tex.stackexchange.com/a/32463/32270
\usepackage[labelfont=bf]{caption}
%% H/T: https://tex.stackexchange.com/a/71153/32270
\usepackage[nottoc,notlot,notlof]{tocbibind}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]

@@

Next, we recall\footnote{We have used slightly non-standard notation for the
terms produced by the de Casteljau algorithm: we start the superscript at
\(n\) and count down to \(0\) as is typically done when describing Horner's
algorithm. For example, we use \(b_j^{(n - 2)}\) instead of
\(b_j^{(2)}\).} the de Casteljau algorithm:

@@

The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
\cite{Mainar1999}, \cite{Farouki1987}):
\begin{equation}
\cond{p, s} = \frac{\widetilde{p}(s)}{
  \left|p(s)\right|} = \frac{\sum_j \left|b_j\right| B_{j, n}(s)}{
  \left|p(s)\right|},
\end{equation}
where \(B_{j, n} \geq 0\).

@@

\noindent  When comparing this computed error to the exact error, the
difference depends only on \(s\) and the Bernstein
coefficients of \(p\). Using a bound (Lemma~\ref{lemma:db-lemma}) on the
round-off error when computing \(\db{1}^{(0)}\), the algorithm can
be shown to be as accurate as if the computations were done in twice
the working precision:

@@

In order to raise from twice the working precision to \(K\) times the
working precision, we continue using EFTs when computing
\(\cdb{1}^{(k)}\). By tracking the round-off from each
floating point evaluation via an EFT, we can form a cascade of global errors:
\begin{align}
  b_j^{(k)} &= \widehat{b}_j^{(k)} + \db{1}_j^{(k)} \\
  \db{1}_j^{(k)} &= \cdb{1}_j^{(k)} + \db{2}_j^{(k)} \\
  \db{2}_j^{(k)} &= \cdb{2}_j^{(k)} +
  \db{3}_j^{(k)} \\
  %% H/T: https://tex.stackexchange.com/a/7651/32270
  &\mathrel{\makebox[\widthof{=}]{\vdots}} \nonumber
\end{align}
In the same way local error can be tracked when updating
\(\widehat{b}_j^{(k)}\), it can be tracked for updates that happen down
the cascade:
\begin{alignat}{4}
  (1 - s) \cdot \widehat{b}_j^{(k + 1)} &+
  s \cdot \widehat{b}_{j + 1}^{(k + 1)} &&  &&=
  \widehat{b}_j^{(k)} &&+ \ell_{1, j}^{(k)} \\
  (1 - s) \cdot \cdb{1}_j^{(k + 1)} &+
  s \cdot \cdb{1}_{j + 1}^{(k + 1)} &&+ \ell_{1, j}^{(k)} &&=
  \cdb{1}_j^{(k)} &&+ \ell_{2, j}^{(k)} \\
  (1 - s) \cdot \cdb{2}_j^{(k + 1)} &+
  s \cdot \cdb{2}_{j + 1}^{(k + 1)} &&+ \ell_{2, j}^{(k)} &&=
  \cdb{2}_j^{(k)} &&+ \ell_{3, j}^{(k)} \\
  &  &&  &&\mathrel{\makebox[\widthof{=}]{\vdots}} && \nonumber
\end{alignat}

@@

\noindent Noting that \(\ell_{F, j}\) contains \(5F - 1\) terms, one can
show that \texttt{CompDeCasteljauK} (Algorithm~\ref{alg:k-comp-de-casteljau})
requires
\begin{equation}
(15K^2 + 11K - 34)T_n + K + 5 =
\mathcal{O}\left(n^2 K^2\right)
\end{equation}
flops to evaluate a degree \(n\) polynomial, where \(T_n\) is the
\(n\)th triangular number. As a comparison, the non-compensated form of
de Casteljau requires \(3 T_n + 1\) flops. In total this will require
\((3K - 4)T_n\) uses of \texttt{TwoProd}. On hardware that supports
FMA, \texttt{TwoProdFMA} (Algorithm~\ref{alg:two-prod-fma}) can be used
instead, lowering the flop count by \(15(3K - 4)T_n\).

@@

\begin{lemma}[\cite{Jiang2010}]
For \(p(s) = \sum_{j = 0}^n b_j B_{j, n}(s)\) a polynomial in Bernstein form
and \(s \in \left[0, 1\right]\), the terms produced by the de Casteljau
algorithm in exact arithmetic satisfy:
\begin{equation}
  \left|b_0^{(0)}\right| \leq \cdots \leq
  \sum_{j = 0}^k \left|b_j^{(k)}\right| B_{j, k}(s) \leq \cdots \leq
  \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s) = \widetilde{p}(s).
\end{equation}
\end{lemma}

\begin{lemma}[\cite{Jiang2010}]\label{lemma:db-lemma}
The second order error \(\db{2}^{(0)}_0\) satisfies
\begin{equation}
  \left|\db{1}^{(0)}_0 - \cdb{1}^{(0)}_0\right| =
  \left|\db{2}^{(0)}_0\right| \leq 2 \gamma_{3n + 1} \gamma_{3(n - 1)}
  \widetilde{p}(s).
\end{equation}
\end{lemma}

@@

Figure~\ref{fig:compensated-k} shows the relative forward errors compared
against the condition number. To compute relative errors, each input and
coefficient is converted to a fraction (i.e. infinite precision) and then
compared to the corresponding computed values. Similar tools are used to
\textbf{exactly} compute the condition number, though here we can rely
on the fact that \(\widetilde{p}(s) = (s - 1)
\left(\frac{s}{2} - \frac{3}{4}\right)^7\). Once the relative errors and
condition numbers are computed as fractions, they are rounded to the
nearest IEEE-754 double precision value. As in \cite{Jiang2010}, we use
values \(\left\{\frac{3}{4} - (1.3)^j\right\}_{j=-5}^{-90}\)\footnote{As with
\(0.1\), it's worth noting that \((1.3)^j\) can't be represented exactly in
IEEE-754 double precision. However, this geometric series still serves a
useful purpose since it continues to raise \(\cond{p, s}\) as \(j\) decreases
away from \(0\) and because it results in ``random'' changes in the bits of
\(0.75\) that are impacted by subtracting \((1.3)^j\).}. The curves for
\texttt{DeCasteljau} and \texttt{CompDeCasteljau} trace the same paths seen
in \cite{Jiang2010}. In particular, \texttt{CompDeCasteljau} has a relative
error that is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
\(1 / \mach\), at which point the relative error increases linearly with
the condition number until it becomes \(\mathcal{O}(1)\) when
\(\cond{p, s}\) reaches \(1 / \mach^2\).
Similarly, the relative error in \texttt{CompDeCasteljau3}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 3\))
is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
\(1 / \mach^2\) at which point the relative error increases linearly
to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^3\)
and the relative error in \texttt{CompDeCasteljau4}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 4\))
is \(\mathcal{O}(\mach)\) until \(\cond{p, s}\) reaches
\(1 / \mach^3\) at which point the relative error increases linearly
to \(\mathcal{O}(1)\) when \(\cond{p, s}\) reaches \(1 / \mach^4\).
